---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# solareclipser

<!-- badges: start -->
<!-- badges: end -->

`solareclipser` is an R package for [SOLAR-Eclipse](https://www.nitrc.org/projects/se_linux/).
It ports SOLAR's Fast Permutation Heritability Inference (FPHI) to R:

- load an empirical kinship pedigree (`IDA`, `IDB`, `KIN`) and a phenotype file
- adjust a trait for covariates, in SOLAR's `covariate` syntax
  (`age^1,2#sex`)
- estimate its heritability (h2r) with standard errors and a likelihood ratio
  test, returned as data frames
- optionally write the results as csv, tsv, json, yaml, or the summary block
  of the original SOLAR `fphi` command

Results are checked against SOLAR 9.0.1 in the package tests.

## Installation

base R installation:

1. Download the tarball from the [releases page](https://github.com/txbrain-org/solareclipser/tree/main/release)
2. Install the package with the following command (replace the path with the
   actual path to the downloaded file):
    ```{r installation, eval = FALSE}
    install.packages("path/to/package.tar.gz", repos = NULL, type = "source")
    ```

## Example

The package bundles a Human Connectome Project kinship pedigree and white
matter phenotypes. The functions take file names, as SOLAR does, so write
them to CSV first:

```{r example}
library(solareclipser)

dir <- tempfile("fphi_")
dir.create(dir)
write.csv(pedigree, file.path(dir, "ped.csv"), row.names = FALSE, quote = FALSE)
write.csv(phenotypes, file.path(dir, "phen.csv"), row.names = FALSE, quote = FALSE)

solar_load_pedigree(file.path(dir, "ped.csv"), threshold = 0, output_dir = dir)
solar_load_phenotype(file.path(dir, "phen.csv"))
solar_select_trait("CC")
res <- solar_run_fphi(file.path(dir, "CC"))
res$results
res$parameters
```

To adjust for covariates, select them before running FPHI, e.g.
`solar_select_covariates("age^1,2#sex")` for age, age^2, sex and their
interactions (the bundled phenotypes have no covariate columns). To also get
the results as text, pass `format`, e.g. `solar_run_fphi(base, format = "solar")`
for output that can be compared with the original SOLAR.

