---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# solareclipser

<!-- badges: start -->
<!-- badges: end -->

## Description

`solareclipser` is an R package to interface SOLAR and to run its main models: polygenic, fphi.

## Installation

You can install the development version of solareclipser like so:

``` r
library(devtools)
install_github("enigma-1590c46634/solareclipser")
```

## Minimal example

```{r polygenic}
library(solareclipser)

sc <- SolarCommand$new()
sc$load(obj = "pedigree",
        fpath = "tests/input/solar/HCP_imputed_filtered_ped.csv",
        cond = "-t 0")
sc$load(obj = "phenotypes",
        fpath = "tests/input/solar/HCP_WM_ave_norm.csv")
sc$trait("CC")$polygenic()
sc$run()
```

```{r fphi}
library(solareclipser)
sc <- SolarCommand$new()
sc$load(obj = "pedigree",
        fpath = "tests/input/solar/HCP_imputed_filtered_ped.csv",
        cond = "-t 0")
sc$load(obj = "phenotypes",
        fpath = "tests/input/solar/HCP_WM_ave_norm.csv")
sc$trait("CC")
sc$create_evd_data(output_fbasename = "CC_evd")
sc$fphi(evd_data = "CC_evd")
sc$run()
```

```{r eval = FALSE}
SolarCommand$new(save_output_dir = "path/to/dir") # to save output files
```

## Tutorial

See [tutorial](inst/doc/tutorial.md) for more infomation.

## SOLAR references

* The new [SOLAR web page](https://solar-eclipse-genetics.org/) (SOLAR-Eclipse)
* [Appendix 1. SOLAR Command Descriptions](http://helix.nih.gov/Documentation/solar-6.6.2-doc/91.appendix_1_text.html)
