---
title: "solareclipser"
output: 
  pdf_document:
    toc: true
  html_vignette:
    toc: true
    toc-location: left
vignette: >
  %\VignetteIndexEntry{solareclipser}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Introduction

`solareclipser` is an R package to interface with [SOLAR-Eclipse](https://www.nitrc.org/projects/se_linux/).

## Installation

### Prerequisites

`solareclipser` requires [`solar`](https://www.nitrc.org/projects/se_linux/).

### solareclipser

```{r installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("txbrain-org/solareclipser")
```

## Usage

`solareclipser` defaults to using a tmp directory for the output files. However,
this can be overridden by passing settings to the `Solar` class.

### `Solar` S6 class

Set the settings for `Solar` class.

```{r settings}
library(solareclipser)

settings <- list(
  output = list(
    dir = "tests/output/solar",
    tcl = FALSE,
    stdout_and_stderr = FALSE
  )
)
```

Note: setting `dir` to a directory from a previous run will allow not needing to load the pedigree and phenotypes again.

### Models

#### Polygenic

```{r polygenic}
solar <- Solar$new(settings = settings)
solar$cmd$load(obj = "pedigree",
               fpath = "tests/input/solar/HCP_imputed_filtered_ped.csv",
               cond = "-t 0")
solar$cmd$load(obj = "phenotypes",
               fpath = "tests/input/solar/HCP_WM_ave_norm.csv")
solar$cmd$trait("CC")$polygenic()
solar$run()
```

#### FPHI

```{r fphi}
library(solareclipser)

solar <- Solar$new(settings = settings)
solar$cmd$load(obj = "pedigree",
               fpath = "tests/input/solar/HCP_imputed_filtered_ped.csv",
               cond = "-t 0")
solar$cmd$load(obj = "phenotypes",
               fpath = "tests/input/solar/HCP_WM_ave_norm.csv")
solar$cmd$trait("CC")
solar$cmd$create_evd_data(output_fbasename = "evd_data")
solar$cmd$fphi(evd_data = "evd_data")
solar$run()
```

### Output

#### Using `SolarFiles`

`SolarFiles` class can be used to get all/individual output files.

```{r solarfiles_polygenic}
# trait_d => trait directory
sf <- SolarFiles$new(settings = settings)
sf$set_polygenic_trait_d(trait_d = "CC")
sfc <- SolarFilesController$new(sf = sf)
sfc$set_polygenic_trait_d_files()
trait_d_files <- sfc$get_polygenic_trait_d_files()

mod_files <- sfc$get_mod_files()
out_files <- sfc$get_out_files()
stat_files <- sfc$get_stats_files()

summary <- readLines(out_files[str_detect(out_files, "polygenic.out")])
cat(str_trim(summary), sep = "\n")
```

#### Using `settings`

```{r settings_tcl_true_stdout_and_stderr_true}
library(solareclipser)

settings <- list(
  output = list(
    dir = "tests/output/solar",
    tcl = TRUE,
    stdout_and_stderr = TRUE
  )
)

solar <- Solar$new(settings = settings)
solar$cmd$load(obj = "pedigree",
               fpath = "tests/input/solar/HCP_imputed_filtered_ped.csv",
               cond = "-t 0")
solar$cmd$load(obj = "phenotypes",
               fpath = "tests/input/solar/HCP_WM_ave_norm.csv")
solar$cmd$trait("CC")$polygenic()
solar$run()
```

## Additional References

- [education - solar-eclipse-genetics.org](https://solar-eclipse-genetics.org/education)
- [solareclipse.pdf - solar-eclipse-genetics.org](https://solar-eclipse-genetics.org/downloads/solareclipser.pdf)
- [www.nitrc.org - solar documents](https://www.nitrc.org/docman/?group_id=558)
